\ \\ \LARGE\bf Speculative Evaluation in Particle Swarm Optimization

### 2010-04-20 honors_thesis/ppsn_paper/paper.tex
removed compactness for compactness

In order to incorporate the determination of which case occurs into the
position and velocity update equations, we introduce an indicator function
$\indic_{t+1}^{\casegen}$ for each case $\casegen \in \caseset$.  When
$\casegen$ corresponds to the case taken by PSO,
$\indic_{t+1}^{\casegen}$ evaluates to 1; otherwise it evaluates to 0.  We can
then sum over all of the cases, and the indicator function will make all of the
terms drop to zero except for the case that actually occurs.  For example, the
indicator function for the specific case $\casexn$ can be written as follows:
$f(x)$
\begin{align}
  \nonumber
	&\indic_{t+1}^{\casexn}\parens{
	  \ofunc\parens{\ppos_{t}},
	  \ofunc\parens{\leftn_{t}},
	  \ofunc\parens{\rightn_{t}},
	  \ofunc\parens{\pbest_{t-1}},
	  \ofunc\parens{\nbest_{t-1}}
	  } \\
  &\hspace{15pt}= \label{eq:deficasexn}
	\begin{cases}
	  1 & \text{if}~
		\ofunc\parens{\ppos_{t}} < \ofunc\parens{\pbest_{t-1}}
	    \,\text{and}\,
		\ofunc\parens{\nbest_{t-1}} < \min\parens{
		  \ofunc\parens{\ppos_{t}},
		  \ofunc\parens{\leftn_{t}},
		  \ofunc\parens{\rightn_{t}}
		} \\
	   0 & \text{otherwise}
	\end{cases}
\end{align}

@@

In the same way we can create notation for the position update function by
substituting into~\eqref{eq:posupdate}.  For compactness, we will drop the
parameters to $\specvel_{t+1}^{\casegen}$ since they can be inferred from the
subscripts.
\begin{align}
\label{eq:defpcasegen}
	\specpos_{t+1}^{\casegen} & \parens{\ppos_{t}, \pvel_{t}, \leftn_{t},
	\rightn_{t} ,\pbest_{t-1} ,\nbest_{t-1}, \pURand_{t}, \nURand_{t}}
	 = \ppos_{t} + \specvel_{t+1}^{\casegen}
\end{align}

With this notation we can re-write the original PSO velocity
equation~\eqref{eq:velupdate}, introducing our sum over cases with the
indicator functions.  Again, we represent the indicator functions and velocity
functions without the parameters for compactness.
The velocity update from~\eqref{eq:velupdate} becomes:
\begin{align}
\label{eq:vel2update}
%\nonumber
	\pvel_{t+1} 
%		\constriction \bigl[ \pvel_t
%			+ \pcoeff\pURand_{t}\otimes\parens{\pbest_{t} - \ppos_{t}}
%			+ \ncoeff\nURand_{t}\otimes\parens{\nbest_{t} -
%			\ppos_{t}} \bigr] \\
%\nonumber
%	&= \sum_{c \in \caseset} \indic_{t+1}^{c} \ \constriction \bigl[ \pvel_t
%			+ \pcoeff\pURand_{t}\otimes\parens{\pbest_{t} - \ppos_{t}}
%			+ \ncoeff\nURand_{t}\otimes\parens{\nbest_{t} -
%			\ppos_{t}} \bigr]  \\
	&= \sum_{c \in \caseset} \ \indic_{t+1}^{c} \ \specvel_{t+1}^{c} 
\end{align}

Similarly, the position update from~\eqref{eq:posupdate}
can also be rewritten using the indicator functions, now with the parameters reinserted:
\begin{align}
\nonumber
  \ppos_{t+1} &= \sum_{c \in \caseset} \bigl[
	\indic_{t+1}^{c}\parens{
	  \ofunc\parens{\ppos_{t}},
	  \ofunc\parens{\leftn_{t}},
	  \ofunc\parens{\rightn_{t}},
	  \ofunc\parens{\pbest_{t-1}},
	  \ofunc\parens{\nbest_{t-1}}
	  } \\
\label{eq:val2updatelong}
	& \quad \quad \quad \specpos_{t+1}^{c}\parens{
	  \ppos_{t},
	  \pvel_{t},
	  \leftn_{t},
	  \rightn_{t},
	  \pbest_{t-1},
	  \nbest_{t-1},
	  \pURand_{t},
	  \nURand_{t}
	  }\bigr]
\end{align}

@@

\begin{algorithm}
  \caption{Speculative Evaluation PSO (SEPSO)}
  \label{alg:centralized}
  \begin{algorithmic}[1]
	\STATE Move all $\p_{t-1}$ to $\noeval{\p}_t$ using
	  \eqref{eq:velupdate}~and~\eqref{eq:posupdate}
	\STATE For each $\noeval{\p}_t$, get its neighbors $\noeval{\nset}_t$ and
	  generate $\noeval{\sset}_{t+1}$ according to
	  \eqref{eq:defpcasegen}.
	\STATE Evaluate all $\noeval{\p}_t$ and $\noeval{\sset}_{t+1}$ in parallel
	\STATE Update personal best for each $\noeval{\p}_t$ and
	  $\noeval{\s}_{t+1}$, creating $\nonbest{\p}_t$ and $\nonbest{\s}_{t+1}$
	\STATE Update neighborhood best for each $\nonbest{\p}_t$, creating
	  $\pset_t$
	\FORALL{$\p_t$}
	\STATE Pick $\nonbest{\s}_{t+1}$ from $\nonbest{\sset}_{t+1}$ that matches
	  the branch taken by $\p_t$ according to
	  \eqref{eq:val2updatelong}.
	\STATE Pass along personal and neighborhood best values obtained by $\p_t$,
	  making $\nonbest{\p}_{t+1}$
	\ENDFOR
	\STATE Update neighborhood best for each $\nonbest{\p}_{t+1}$, creating
	  $\pset_{t+1}$
	\STATE Repeat from Step 1 until finished
  \end{algorithmic}
\end{algorithm}

>>>>

In order to incorporate the determination of which case occurs into the
position and velocity update equations, we introduce an indicator function
$\indic_{t+1}^{\casegen}$ for each case $\casegen \in \caseset$.  When
$\casegen$ corresponds to the case taken by PSO,
$\indic_{t+1}^{\casegen}$ evaluates to 1; otherwise it evaluates to 0.  We can
then sum over all of the cases, and the indicator function will make all of the
terms drop to zero except for the case that actually occurs.  For example, the
indicator function for the specific case $\casexn$ can be written as follows:
\begin{align}
  \nonumber
	&\indic_{t+1}^{\casexn}\parens{
	  \ofunc\parens{\ppos_{t}},
	  \ofunc\parens{\leftn_{t}},
	  \ofunc\parens{\rightn_{t}},
	  \ofunc\parens{\pbest_{t-1}},
	  \ofunc\parens{\nbest_{t-1}}
	  } \\
  &\hspace{15pt}= \label{eq:deficasexn}
	\begin{cases}
	  1 & \text{if}~
		\ofunc\parens{\ppos_{t}} < \ofunc\parens{\pbest_{t-1}}
	    \,\text{and}\,
		\ofunc\parens{\nbest_{t-1}} < \min\parens{
		  \ofunc\parens{\ppos_{t}},
		  \ofunc\parens{\leftn_{t}},
		  \ofunc\parens{\rightn_{t}}
		} \\
	   0 & \text{otherwise}
	\end{cases}
\end{align}

@@

In the same way we can create notation for the position update function by
substituting into~\eqref{eq:posupdate}:
\begin{align}
    \nonumber
	\specpos_{t+1}^{\casegen} & \parens{\ppos_{t}, \pvel_{t}, \leftn_{t},
	\rightn_{t} ,\pbest_{t-1} ,\nbest_{t-1}, \pURand_{t}, \nURand_{t}}
        \\
    \label{eq:defpcasegen}
        &=
	\ppos_{t} + \specvel_{t+1}^{\casegen}
	\parens{\pvel_t, \ppos_{t}, \leftn_{t}, \rightn_{t},
	\pbest_{t-1}, \nbest_{t-1}, \pURand_{t}, \nURand_{t}}
\end{align}

With this notation we can re-write the original PSO velocity
equation~\eqref{eq:velupdate}, introducing our sum over cases with the
indicator functions.
The velocity~\eqref{eq:velupdate} and position~\eqref{eq:posupdate} equations become:
\begin{align}
\nonumber
  \pvel_{t+1} &= \sum_{c \in \caseset} \bigl[
	\indic_{t+1}^{c}\parens{
	  \ofunc\parens{\ppos_{t}},
	  \ofunc\parens{\leftn_{t}},
	  \ofunc\parens{\rightn_{t}},
	  \ofunc\parens{\pbest_{t-1}},
	  \ofunc\parens{\nbest_{t-1}}
	  } \\
\label{eq:specvelupdate}
	& \quad \quad \quad \specvel_{t+1}^{c}\parens{
	  \ppos_{t},
	  \pvel_{t},
	  \leftn_{t},
	  \rightn_{t},
	  \pbest_{t-1},
	  \nbest_{t-1},
	  \pURand_{t},
	  \nURand_{t}
	  }\bigr]
\\
\nonumber
  \ppos_{t+1} &= \sum_{c \in \caseset} \bigl[
	\indic_{t+1}^{c}\parens{
	  \ofunc\parens{\ppos_{t}},
	  \ofunc\parens{\leftn_{t}},
	  \ofunc\parens{\rightn_{t}},
	  \ofunc\parens{\pbest_{t-1}},
	  \ofunc\parens{\nbest_{t-1}}
	  } \\
\label{eq:specposupdate}
	& \quad \quad \quad \specpos_{t+1}^{c}\parens{
	  \ppos_{t},
	  \pvel_{t},
	  \leftn_{t},
	  \rightn_{t},
	  \pbest_{t-1},
	  \nbest_{t-1},
	  \pURand_{t},
	  \nURand_{t}
	  }\bigr]
\end{align}

@@

\begin{algorithm}
  \caption{Speculative Evaluation PSO (SEPSO)}
  \label{alg:centralized}
  \begin{algorithmic}[1]
	\STATE Move all $\p_{t-1}$ to $\noeval{\p}_t$ using
	  \eqref{eq:velupdate}~and~\eqref{eq:posupdate}
	\STATE For each $\noeval{\p}_t$, get its neighbors $\noeval{\nset}_t$ and
	  generate $\noeval{\sset}_{t+1}$ according to
	  \eqref{eq:defpcasegen}.
	\STATE Evaluate all $\noeval{\p}_t$ and $\noeval{\sset}_{t+1}$ in parallel
	\STATE Update personal best for each $\noeval{\p}_t$ and
	  $\noeval{\s}_{t+1}$, creating $\nonbest{\p}_t$ and $\nonbest{\s}_{t+1}$
	\STATE Update neighborhood best for each $\nonbest{\p}_t$, creating
	  $\pset_t$
	\FORALL{$\p_t$}
	\STATE Pick $\nonbest{\s}_{t+1}$ from $\nonbest{\sset}_{t+1}$ that matches
	  the branch taken by $\p_t$ according to
	  \eqref{eq:specposupdate}.
	\STATE Pass along personal and neighborhood best values obtained by $\p_t$,
	  making $\nonbest{\p}_{t+1}$
	\ENDFOR
	\STATE Update neighborhood best for each $\nonbest{\p}_{t+1}$, creating
	  $\pset_{t+1}$
	\STATE Repeat from Step 1 until finished
  \end{algorithmic}
\end{algorithm}
